Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-05-29 doc/network.tex
trying to move things around...

Here, we investigate two methods for correcting for temporal bias in a contact tracing data.
We refer to the first method as the population-level method as it relies on the observed distribution aggregated across the entire population.
As the observed generation-interval distribution is a weighted distribution of the effective distribution, the effective distribution can be recovered by taking the inverse of the weights, given by \eref{exp}.
Hence, the population-level method relies on the observed generation intervals and the exponential growth rate \citep{tomba2010some, nishiura2010time, britton2019estimation}:
\begin{equation}
g(\tau) \propto \tsub{c}{exp}(\tau) \exp(r\tau).
\end{equation}
When the susceptible dynamics is known, this equation can be generalized by using \eref{obsg}; we do not pursue this direction here.
The population-level method is simple and intuitive, but does not take into account who infected whom.

We refer to the second method as the individual-level method because it relies on individual contact information.
We model each infection $i$ from an infected individual $j$ as a non-homogeneous Poisson process between the time at which infector $j$ was infected ($t_j$) and the censorship time ($\tsub{t}{censor}$), where the time-varying Poisson rate at time $t$ is equal to $\RR g(t - t_j)$, where $g(t)$ is the effective generation-interval distribution \citep{daley2007introduction}.
\dc{is effective GI distribution, $g$, the same as the intrinsic? The notation implies it is... but I'm confused now.}
We use a gamma distribution (parameterized by its mean and shape) to model the effective generation-interval distribution.
Then, the probability that an individual $j$ infects $n_j$ individuals between $t_j$ and $\tsub{t}{censor}$ is equal to
\begin{equation}
\frac{\RR^{n_j} \exp \big(- \RR G(\tsub{t}{censor} - t_j ;\theta) \big) \prod_{i=1}^{n_j} g(s_{i, j}; \theta)}{n_j!},
\end{equation}
where $s_{i,j}$ is the observed generation interval between infector $j$ and infectee $i$, and $\theta$ is a (vector) parameter of the generation-interval distribution $g$ (and the corresponding cumulative distribution function $G$).

To test these methods, we simulate 100 epidemics with Ebola-like parameters on an empirical network \citep{leskovec2016snap} and compare the estimates of reproductive number with empirical reproductive numbers, which we define as the average number of secondary cases generated by the first 100 infected individuals (\fref{cmp}).
\cite{trapman2016inferring} defined the empirical reproductive number by dividing the total number of infected individuals in generations $2$ to $k+1$ by the total number of infected individuals in generations $1$ to $k$, where the reference generation $k$ is defined to be the first generation that contains at least 75 infected individuals;
however, individuals in later generations are potentially subject to stronger local depletion and even global depletion effects.
Therefore, we expect our empirical estimate to better reflect the infection process during the exponential growth phase.
\swp{TODO: Compare Trapman vs our estimate of empirical R0 in Appendix; Trapman estimate is much lower.}

@@

>>>>

Here, we investigate two methods for correcting for temporal bias in a contact tracing data (see Methods for details).
We refer to the first method as the population-level method as it relies on the observed distribution aggregated across the entire population.
When an epidemic is growing exponentially, the right-censoring causes the observed generation-interval inverse exponentially weighted \eref{exp}; hence, we can ``undo'' the censoring by exponentially weighting the observed generation-interval distribution \citep{tomba2010some, nishiura2010time, britton2019estimation}:
\begin{equation}
\tsup{g}{exp}(\tau) \propto c_t(\tau) \exp(r\tau),
\end{equation}
where $r$ is the exponential growth rate.
We refer to the second method as the individual-level method because it relies on individual contact information.
We model each infection as a non-homogeneous Poisson process arising from the infector; 
incorporating information about time of infection of an infector, time of infection of an infectee, and time since the beginning of an epidemic allows us to explicitly model the censoring process in the observed generation intervals.

To test these methods, we simulate 100 epidemics with Ebola-like parameters on an empirical network \citep{leskovec2016snap} and compare the estimates of reproductive number with empirical reproductive numbers, which we define as the average number of secondary cases generated by the first 100 infected individuals (\fref{cmp}).
\cite{trapman2016inferring} defined the empirical reproductive number by dividing the total number of infected individuals in generations $2$ to $k+1$ by the total number of infected individuals in generations $1$ to $k$, where the reference generation $k$ is defined to be the first generation that contains at least 75 infected individuals.
however, individuals in later generations are potentially subject to stronger local depletion and even global depletion effects.
Therefore, we expect our empirical estimate to better reflect the infection process during the exponential growth phase.
\swp{TODO: Compare Trapman vs our estimate of empirical R0 in Appendix; Trapman estimate is much lower.}

@@

\swp{TODO:}
\subsection{Egocentric ...}

\subsection{Population-level method}

\subsection{Individual-level method}

We model each infection $i$ from an infected individual $j$ as a non-homogeneous Poisson process between the time at which infector $j$ was infected ($t_j$) and the censorship time ($\tsub{t}{censor}$), where the time-varying Poisson rate at time $t$ is equal to $\RR g(t - t_j)$, where $g(t)$ is the effective generation-interval distribution \citep{daley2007introduction}.
\dc{is effective GI distribution, $g$, the same as the intrinsic? The notation implies it is... but I'm confused now.}
We use a gamma distribution (parameterized by its mean and shape) to model the effective generation-interval distribution.
Then, the probability that an individual $j$ infects $n_j$ individuals between $t_j$ and $\tsub{t}{censor}$ is equal to
\begin{equation}
\frac{\RR^{n_j} \exp \big(- \RR G(\tsub{t}{censor} - t_j ;\theta) \big) \prod_{i=1}^{n_j} g(s_{i, j}; \theta)}{n_j!},
\end{equation}
where $s_{i,j}$ is the observed generation interval between infector $j$ and infectee $i$, and $\theta$ is a (vector) parameter of the generation-interval distribution $g$ (and the corresponding cumulative distribution function $G$).
